---
title: "Data Wrangling Training"
author: "Iwen Su"
date: "3/6/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Intro to Data Wrangling: `dplyr`

### Wrangle gapminder data using `dplyr`.

First load the packages.

```{r load, message=FALSE}

library(tidyverse)
library(skimr)

```

### Get data from github.

```{r data, message=FALSE}

gapminder <- readr::read_csv("https://raw.githubusercontent.com/OHI-Science/data-science-training/master/data/gapminder.csv")

```

### Explore gapminder data.

```{r explore}

head(gapminder)
tail(gapminder)

head(gapminder, 10) # see the first ten rows
tail(gapminder, 12) # see the last twelve rows

## See structure and other info about the data frame
str(gapminder)
names(gapminder) 
dim(gapminder) 
c(nrow(gapminder), ncol(gapminder)) # create dim() from nrow() and ncol()

## summary statistics
summary(gapminder)
skim(gapminder) 


```

> Note: In addition to what data.frames can store, tibble can store lists, ggplots, and multiple dataframes.

Explore inside the dataframe `$`

```{r}

head(gapminder$lifeExp, 10)
str(gapminder$lifeExp) # more useful function for looking at entire dataset
summary(gapminder$lifeExp)
skim(gapminder$lifeExp)

```

### Dplyr basics: 

Main functions include: `filter()`, `select()`, `mutate()`, `summarise()`, and `arrange()`. See dplyr [cheatsheet](http://www.rstudio.com/wp-content/uploads/2015/02/data-wrangling-cheatsheet.pdf).

```{r}

filter(gapminder, lifeExp < 29)

# Filter for 1 country
filter(gapminder, country == "Mexico")

# Filter for multiple countries - use the `%in%` operator on `c()`
filter(gapminder, country %in% c("Mexico", "Chile"))

# Filter for 1 country in a specific year
filter(gapminder, country == "Mexico", year == 2002)

# Select by columns
select(gapminder, year, lifeExp)

# Use the `-` to deselect columns
select(gapminder, -continent, -lifeExp)

# Use filter and select together
gap_cambodia <- filter(gapminder, country == "Cambodia") 
gap_cambodia2 <- select(gap_cambodia, -lifeExp, -continent)

```

### Life changing pipe operator `%>%` (and then)

```{r}

gapminder %>% head() # equivalent to head(gapminder)

gapminder %>% head(3) # equivalent to head(gapminder, 3)

# Use filter and select together 
gap_cambodia <- filter(gapminder, country == "Cambodia") 
gap_cambodia2 <- select(gap_cambodia, -lifeExp, -continent)

# now let's do that with the %>% operator
gap_cambodia <- gapminder %>% 
  filter(country == "Cambodia")

gap_cambodia2 <- gap_cambodia %>% 
  select(-lifeExp, -continent)

# or you could combine the two variables created above by stacking `%>%` operators
gap_cambodia3 <- gapminder %>% 
  filter(country == "Cambodia") %>% 
  select(-continent, -lifeExp)

```

### To add variables use `mutate()`

* You can run other functions inside the `mutate()` function
* You can also run mathematical operations on existing columns

```{r}

gapminder %>% 
  mutate(index = 1:nrow(gapminder))

gapminder %>% 
  mutate(gdp = pop*gdpPercap) # can't call 'pop' or 'gdpPercap' by itself

```

### Exercise: Find the maximum gdpPercap of Egypt and Vietnam, and save it in a new column

1. First filter for Egypt and Vietnam
2. Then use `mutate` to create a new column that has the max GDP between Egypt and Vietnam together
3. What if we want the max GDP of Egypt and max GDP of Vietnam separately? Use `group_by`

```{r}

# Find max GDP per capita of both Egypt and Vietnam
gapminder %>% 
  filter(country %in% c("Egypt","Vietnam")) %>% 
  mutate(max_gdp = max(gdpPercap))

# Find max GDP per capita of Egypt and Vietnam separately
x2 <- gapminder %>% 
  filter(country %in% c("Egypt","Vietnam")) %>% 
  group_by(country) %>% 
  mutate(max_gdp = max(gdpPercap)) %>% 
  ungroup() # important to remember to ungroup

str(x2)

```

### Summarize and `group_by` together

While it can be nice to replicate the `max_gdp` value for each observation especially when you are running commands on all the observations, you may also want to just select the single value of max gdp for each country.

> Note: 'CTRL+I' is the shortcut for reindenting lines.

```{r}

gapminder %>% 
  group_by(country) %>% 
  mutate(gdp = pop*gdpPercap) %>% 
  summarize(max_gdp = max(gdp)) %>% 
  ungroup() %>% 
  arrange(max_gdp) # ascending order

gapminder %>% 
  group_by(country) %>% 
  mutate(gdp = pop*gdpPercap) %>% 
  summarize(max_gdp = max(gdp)) %>% 
  ungroup() %>% 
  arrange(desc(max_gdp)) # descending order



```

Let's find the max lifeExp for countries in Asia

```{r}

# Grab max GDP for just countries in Asia
asia <- gapminder %>% 
  filter(continent == "Asia") %>% 
  group_by(country) %>% 
  mutate(gdp = pop*gdpPercap) %>% 
  summarize(max_gdp = max(gdp)) %>% 
  ungroup() %>% 
  arrange(desc(max_gdp)) # descending order

# Grab max life expectancy for countries in Asia
gapminder %>% 
  filter(continent == "Asia") %>% 
  group_by(country) %>% 
  summarize(max_gdp = max(lifeExp)) %>% 
  ungroup() %>% 
  dplyr::arrange(desc(max_gdp)) # descending order


```


### Joining Datasets

Also referred to as relational data, because there is a relationship between two or more data tables. An example is that the same column header (variable) exists in both data tables.

* You have to decide which data table you want to maintain the observations for, as they may share the same column header, but all the observations may not have match for each observation
* `left_join` and `right_join` can be used to determine whether you want to maintain the integrity of data table on the left (first argument value) or on the right (second argument value) in the function
* `inner_join` keeps only matched rows between both data tables
* `full_join` retains all rows even if there are no matches

```{r join}

# Bring in second data table that will be joined with the `gapminder` dataset
co2 <- read_csv("https://raw.githubusercontent.com/OHI-Science/data-science-training/master/data/co2.csv")

# explore
co2 %>%
  head()

# Select for only 2007 data since co2 data is only from that year
gap2007 <- gapminder %>% 
  filter(year == 2007)

# Left_join gap2007 to co2, by which column?
lj <- left_join(gap2007, co2, by = "country")

# Explore lj
lj %>% dim()
lj %>% summary()

# Right_join gap2007 to co2, by which column?
rj <- right_join(gap2007, co2, by = "country")

# Explore rj
rj %>% dim()
rj %>% summary()

```

<br>
<br>

## Intro to Data Wrangling: `tidyr`

Data wrangling with `tidyr`, which is part of the tidyverse. We are going to tidy some data!

* `gather()`
* `separate()`
* `spread()`
* `

### Load `tidyverse`

```{r message=FALSE}

library(tidyverse) # install.packages("tidyverse")

```

### Read in gapminder dataset wide and long

*ID variables* (key) are those that identify an observation such as continent, country or year.
*Observation variables* (value) are values of the observation such as population, life expectancy and gdp per capita.

```{r message=FALSE, results='hide'}

# wide format
gap_wide <- readr::read_csv("https://raw.githubusercontent.com/OHI-Science/data-science-training/master/data/gapminder_wide.csv")

## long format (intermediate)
gapminder <- readr::read_csv('https://raw.githubusercontent.com/OHI-Science/data-science-training/master/data/gapminder.csv')

```

### Take a look

```{r}

head(gap_wide)
str(gap_wide)

```

### Use `gather()` on the wide format data table
Learned a new function `starts_with()`

```{r}

# attempt 1 - give names to new key-value columns
gap_long <- gap_wide %>% 
  gather(key = obstype_year,
         value = obs_values)

# unfortunately need to specify what you want the gather() function to gather
str(gap_long)
head(gap_long)
tail(gap_long)

# tell gather() which columns you want to apply it to
gap_long <- gap_wide %>% 
  gather(key = obstype_year,
         value = obs_values,
         dplyr::starts_with('pop'),
         dplyr::starts_with('lifeExp'),
         dplyr::starts_with('gdpPercap')
         )

str(gap_long)
head(gap_long)
tail(gap_long)

# you can also use `-` to tell gather() which columns you want to exclude
gap_long <- gap_wide %>% 
  gather(key = obstype_year,
         value = obs_values,
         -continent, -country)

str(gap_long)
head(gap_long)
tail(gap_long)

```

### Using `separate()` to turn one column into multiple columns

```{r}

gap_long <- gap_wide %>% 
  gather(key = obstype_year,
         value = obs_values,
         -continent, -country) %>% 
  separate(obstype_year,
           into = c('obs_type', 'year'),
           sep = "_",
           convert = TRUE) # R converts values in new columns to what it think it should be (integer, numeric, or logical)

# great, `year` column is of class integer
str(gap_long)
head(gap_long)
tail(gap_long)

```

### Check it out in `ggplot`:
Plot Canada's life expectancy

```{r}

canada_df <- gap_long %>% 
  filter(obs_type == 'lifeExp',
         country == 'Canada')

ggplot(canada_df, aes(x = year, y = obs_values)) +
  geom_line()

```


### Check it out in `ggplot`:
Plot life expectancy in all countries in Americas

```{r}

life_df <- gap_long %>% 
  filter(continent == "Americas", obs_type == "lifeExp")

ggplot(life_df, aes(x = year, y = obs_values, color = country)) +
  geom_line()

```

### Exercise

Use `gap_long`:
* calculate and plot the mean life expectancy for each continent over time from 1982 to 2007
* give your plot a title and assign x and y labels

```{r}

life_all <- gap_long %>% 
  filter(year >= 1982, obs_type == 'lifeExp') %>% 
  group_by(continent, year) %>% 
  summarize(meanLifeExp = mean(obs_values)) %>% 
  ungroup()

head(life_all)
tail(life_all)

ggplot(life_all, aes(x = year, y = meanLifeExp, color = continent)) +
  geom_line() +
  labs(x = "Year", 
       y = "Mean Life Expectancy", 
       title = "Mean Life Expectancy across Continents from 1982 to 2007" ) +
  theme_classic()

```


```{r}

```

