---
title: "ggplot2_training"
author: "Iwen Su"
date: "2/23/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load packages

```{r package, message = F}

library(tidyverse)

```

## Get mpg dataset

```{r}

mpg

```

## Plot data as scatterplot

a. geom_point() - creates simple scatterplot

```{r}

ggplot(data = mpg, aes(x = displ, y = hwy)) + # how does mileage compare to engine size
  geom_point()

```

## If creating multiple plots from same data

```{r}

# assign first line of ggplot() to a variable

car_plot <- ggplot(data = mpg, aes(x = displ, y = hwy))

# build off of car_plot - useful when trying to create different visualizations from the same base data

car_plot +
  geom_point()

```

## Adjust transparency of points in scatterplot

```{r}

ggplot(data = mpg, aes(x = displ, y = hwy)) +
  geom_point(alpha = 0.4) # adjusts transparency and reveals that there are multiple dots on the same coordinate point

```

## Group points using the class category by color or by shape

```{r}

ggplot(data = mpg) +
  geom_point(aes(x = displ, y = hwy, color = class))

m <- ggplot(data = mpg, aes(x = displ, y = hwy)) 

m +
  geom_point(aes(color = class))

m +
  geom_point(aes(shape = class)) # there are a limited number of shapes so 'suv' was not assigned a shape

```

## ggplot exercise: plot cty and hwy as scatterblot by car class and fuel type

```{r}

ggplot(data = mpg) +
  geom_point(aes(x = cty, y = hwy, size = class, color = fl)) # group data 'class' by size, 'fl' by color

```

```{r}

ggplot(mpg, aes(displ,hwy))


ggplot(data = mpg) +
  geom_point(aes(x = displ, y = hwy), color = "blue", alpha = 0.4) # when adding arguments outside of aes, it is applying color and transparency to the entire data set (e.g. make all points blue and transparency of 0.4)

```

## More ggplot exercises from data visualization training doc (http://r4ds.had.co.nz/data-visualisation.html)

```{r}

ggplot(data = mpg) +
  geom_point(aes(x = displ, y = hwy, colour = cty)) # applying colour gradient to a continuous variable 'cty'

ggplot(data = mpg) +
  geom_point(aes(x = displ, y = hwy, colour = displ < 5)) # give color to all "displ" values < 5

```

