---
title: "ggplot2_training"
author: "Iwen Su"
date: "2/23/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load packages

```{r package, message = F}

library(ggthemes)
library(tidyverse)

```

## Get mpg dataset

```{r}

mpg

```

## Plot data as scatterplot

a. geom_point() - creates simple scatterplot

```{r}

ggplot(data = mpg, aes(x = displ, y = hwy)) + # how does mileage compare to engine size
  geom_point()

```

## If creating multiple plots from same data

```{r}

# assign first line of ggplot() to a variable

car_plot <- ggplot(data = mpg, aes(x = displ, y = hwy))

# build off of car_plot - useful when trying to create different visualizations from the same base data

car_plot +
  geom_point()

```

## Adjust transparency of points in scatterplot

```{r}

ggplot(data = mpg, aes(x = displ, y = hwy)) +
  geom_point(alpha = 0.4) # adjusts transparency and reveals that there are multiple dots on the same coordinate point

```

## Group points using the class category by color or by shape

```{r}

ggplot(data = mpg) +
  geom_point(aes(x = displ, y = hwy, color = class))

m <- ggplot(data = mpg, aes(x = displ, y = hwy)) 

m +
  geom_point(aes(color = class))

m +
  geom_point(aes(shape = class)) # there are a limited number of shapes so 'suv' was not assigned a shape

```

## ggplot exercise from [training doc][http://r4ds.had.co.nz/data-visualisation.html]: plot cty and hwy as scatterblot by car class and fuel type

```{r}

ggplot(data = mpg) +
  geom_point(aes(x = cty, y = hwy, size = class, color = fl)) # group data 'class' by size, 'fl' by color

```

## specifying arguments outside of aes

```{r}

ggplot(data = mpg) +
  geom_point(aes(x = displ, y = hwy), color = "blue", alpha = 0.4) # when adding arguments outside of aes, it is applying color and transparency to the entire data set (e.g. make all points blue and transparency of 0.4)

```

## More ggplot exercises 

From data visualization [training doc][http://r4ds.had.co.nz/data-visualisation.html]
a. apply colour gradient
b. distinguish between values < 5 and values > 5

```{r}

ggplot(data = mpg) +
  geom_point(aes(x = displ, y = hwy, colour = cty)) # applying colour gradient to a continuous variable 'cty'

ggplot(data = mpg) +
  geom_point(aes(x = displ, y = hwy, colour = displ < 5)) # give color to all "displ" values < 5

```

## Faceting 

```{r}

ggplot(data = mpg) +
  geom_point(aes(x = displ, y = hwy, color = class)) +
  facet_wrap(~ manufacturer) # wrap this data by (~) 'manufacturer

```

## ggplot Themes

See more customized themes by [Jeffrey Arnold][https://cran.r-project.org/web/packages/ggthemes/vignettes/ggthemes.html]. 

```{r}

theme <- ggplot(data = mpg) +
  geom_point(aes(x = displ, y = hwy, color = class)) +
  facet_wrap(~ manufacturer)

theme +
  theme_minimal()

theme +
  theme_wsj()

```

## Looking into other Geoms

```{r}

ggplot(mpg, aes(x = drv, y = hwy)) +
  geom_jitter()

ggplot(mpg, aes(x = drv, y = hwy)) +
  geom_boxplot()

ggplot(mpg, aes(x = drv, y = hwy)) +
  geom_violin() # includes density per hwy value
 
```

```{r}

ggplot(data = mpg) +
  geom_point(aes(x = displ, y = hwy))

ggplot(data = mpg) +
  geom_smooth(aes(x = displ, y = hwy))

# combine scatterplot and smoothing - graph it using two different geoms

ggplot(data = mpg, aes(x = displ, y = hwy)) +
  geom_point() +
  geom_smooth()

```

```{r}

ggplot(data = mpg, aes(x = displ, y = hwy)) +
  geom_point(aes(color = class)) + # only want class grouping (by color) to apply to points
  geom_smooth()

```

### Exercise: Customizing my ggplot

```{r}

ggplot(data = mpg, aes(x = displ, y = hwy)) +
  geom_point(aes(color = class)) + # only want class grouping (by color) to apply to points
  geom_smooth() +
  theme_minimal() +
  labs(title = "Relationship between Engline Size and Miles per Gallon", x = "Engine Displacement (litres)", y = "Highway (mpg)") +
  scale_color_discrete(name = "Type of Car", labels = c("two-seater","compact","midsize","minivan","pickup","subcompact","suv")) +
  theme(legend.position = "bottom")
  

```

Bar charts

```{r}

bar <- ggplot(data = mpg) 

bar +
  geom_bar(aes(x = fl)) # default is to count number of observations in each bin or categorical value in 'fl'

bar +
  geom_bar(aes(x = fl, color = fl)) 

bar +
  geom_bar(aes(x = fl, fill = fl)) # since x and fill arguments are the same, it just fills it with a color

bar +
  geom_bar(aes(x = fl, fill = class)) # bins are from 'fl' and within each bin 'class' groupings are separated

bar +
  geom_bar(aes(x = fl, fill = class), position = "fill") # proportions of class within each bin is represesnted

bar +
  geom_bar(aes(x = fl, fill = class), position = "dodge") #align each 'class' group next to each other rather than stacking

bar +
  geom_bar(aes(x = fl, fill = class), position = "dodge") +
  theme_minimal() +
  scale_x_discrete(labels = c("CNG", "Diesel", "Ethanol", "Premium", "Regular")) +
  labs(x = "Fuel Type", y = "Number of cars", title = "Here is my customized graph") +
  scale_fill_brewer(palette = "PRGn")

# see palettes that can be used in grouping values by color
RColorBrewer::display.brewer.all()  

``` 

# Saving image

```{r}

myplot <- ggplot(data = mpg) +
  geom_bar(aes(x = fl, fill = class), position = "dodge") +
  theme_minimal() +
  scale_x_discrete(labels = c("CNG", "Diesel", "Ethanol", "Premium", "Regular")) +
  labs(x = "Fuel Type", y = "Number of cars", title = "Here is my customized graph") +
  scale_fill_brewer(palette = "PRGn")

ggsave("myplot.png", myplot, width = 6, height = 6)

ggsave("myplot.pdf", myplot, width = 6, height = 6) # smaller file size

```

