---
title: "purrr"
author: "Iwen Su"
date: "7/20/2018"
output: html_document
---

# Exploring `purrr` Functionalities

About: The `purrr` package makes it easy to work with lists and functions. This cheatsheet will remind you how to manipulate lists with purrr as well as how to apply functions iteratively to each element of a list or vector. 

```{r}
library(purrr)
library(repurrrsive)
```

## `purrr` basics

`purrr` guesses what you want to do
`map` is your best friend

```{r}

## Create a list
shades <- colors()[1:10] # the list we want to apply a function to 

## Print each of the 10 colors in `shades`
## option 1: for loop function
for (i in seq_along(shades)){
  print(shades[i])
}

## option 2: using lapply
a <-  lapply(shades, print)

## option 3: using purrr
a <-  map(shades, print)

```

## More about `map`

By default returns a list. Different map functions return different objects

* `map_lgl` - logical objects
* `map_df` - data frames
* `map2` - applies function to two vectors/lists; first two arguments are lists, third is the function `map2(list1, list2, function)`
* `pmap` - applies function to more than 2 vectors/lists; first argument is a list of list, second is the function `pmap(list(list1,list2,list3), function,...)`

```{r}

head(mtcars)

## Calculate mean of each column (mpg, cyl, disp, hp, etc) in mtcars
map_df(mtcars,mean,na.rm=T)

## Compare to check
mean(mtcars$mpg)
mean(mtcars$cyl)

## Pasting two vectors
a <- map2_chr(c('one','two','red','blue'), c('fish'), paste); a
b <- map2(c('one','two','red','blue'), c('fish'), paste); b

fish_colors <- list(c('one','two','red','blue'), c('fish'))
pmap_chr(fish_colors,paste)
class(a)

## Specify exactly how you want to paste the two vectors together
## I think the squiggly indicates you are linking to the first and second argument, w/o it doesn't work
map2_chr(c('one','two','red','blue'), c('fish'), ~paste(.x,.y))
map2_chr(c('one','two','red','blue'), c('fish'), ~paste(.y,.x))
map2_chr(c('one','two','red','blue'), c('fish'), paste(.x,.y))

```

## Perks of `purrr::map`

The cool thing that `purrr` offers is you don't have to `rbind`, `cbind`, or `join` your dataframes into one before passing a function over it as in `lapply`, where you only get to specify a function over a single vector or list.

```{r} lapply(X, FUN..)```, where X is a vector or list

### Let's try with the `got_chars` data.

* Can specific a character vector of names or a numeric vector of position - only these will be modified
* `got_list` only prints out the 4 elements named above
* In this method, you still preserve the entire list of 30, pulling out the 4 elements of interest
```{r}

got_list_test <- got_chars %>%
  map(`[`, c('name','aliases','allegiances','born'))
str(got_list, list.len=4)
```

You can rearrange and group same elements into lists. The `got_chars` list contains 30 lists
```{r}
View(got_chars)
length(got_chars)
```


### Create list for each element from the 30 lists in got_chars
```{r}
k <- map_chr(got_chars,'name')
map(got_chars,'aliases')
map(got_chars,'allegiances')
map_chr(got_chars,'born')
```

Create a list for each of the 4 elements - group them together
```{r}
got_list <-  got_chars %>% {
  list(
    name = map_chr(.,'name'),
    aliases = map(.,'aliases'),
    allegiances = map(.,'allegiances'),
    born = map_chr(.,'born')
  )
}

## Take a look at list structure, 4 lists
str(got_list, list.len=4)


```


Paste the name, aliase, allegiance, and birth information together across the lists. `got_list2[["name"]][1]` matches with `got_list2[["aliases"]][1]`, `got_list2[["allegiances"]][1]` and `got_list2[["born"]][1]`
```{r}

## Create function for getting name,  number of aliases, number of allegiances, and the born information
got_foo <- function(name, aliases, allegiances,born){

  paste(name, 'has', length(aliases), 'aliases and', length(allegiances),
        'allegiances, and was born in', born)

}

got_list %>%
  pmap_chr(got_foo) %>%
  head()


## This doesn't work because got_list contains more than 2 lists
got_list %>%
  map_chr(got_foo) %>%
  head()

```

`purrr` functions to view different parts of a list

```{r}
str(got_list, list.len=4)

## you can just grab the allegiances, but this will turn it back into a list of 30; each allegiance you pluck becomes a list
allegiances <- pluck(got_list,"allegiances"); View(allegiances)
```






***







## Subset for an item across multiple lists

Unlike `lapply`, do not need to specify a function to select for each
```{r}

lapply(got_chars[1:3], 
       function(x) x[["name"]])

map(got_chars[1], "name")

## View b again
b
## Name the elements in the list
names(b) <- c("A","B","C","D")
## Create a list that contains three lists
b_list <- list(b,b,b)
## Select for element A in all three lists
map(b_list[1:2],"A")
map(b[1],"A")
```


## Apply linear regression model for each country using `group_by`

```{r}
d_gf <- d %>%
  dplyr::group_by(country) %>%
  dplyr::do({ 
    mod <- lm(value ~ year, data =.)
    value_pred <- predict(mod, newdata =.[c('year')]) # value_pred = country-grouped mod$fitted.values?
    data.frame(., value_pred) # do loop applies the model fitting and prediction to each country group
    }) %>% 
  dplyr::ungroup()
```

