---
title: "Seaside Chat 8/20"
author: "Iwen Su"
output: html_document
---


# Common String/RegEX Uses in OHI

## OHI Ex: Extract date or number string

1. Setup
```{r}
library(raster)
library(sp)
library(stringr)

source('https://rawgit.com/OHI-Science/ohiprep_v2018/master/src/R/common.R')
npp_files_gf <- list.files(file.path(dir_M,"git-annex/globalprep/prs_fish/v2018/VGPM_primary_productivity/int/annual_npp"), full.names=TRUE, pattern = "gf")
```

2. Extract string of digits - useful in OHI when matching up the data year for different data
```{r}

file <- "/home/shares/ohi/git-annex/globalprep/prs_fish/v2018/int/comm_landings/annual_catch_2003.tif"
str_extract(basename(file),"(\\d)+") # extracts any digits


catch_npp_fun <- function(file, layer){
  
  catch <- raster(file)
  base <- basename(file)
  
  #base <- "annual_2000_2003"
  #base <- "annual_20002003"
  
  yr <- str_extract(base,"(\\d)+") # extracts the first consecutive digit string
  
  # Grab the NPP raster from the same year
  npp <- npp_files_gf[str_detect(npp_files_gf, yr)] %>% 
    raster()

  # Resample and project
  projectRaster(catch, npp, method = 'ngb', over=TRUE, filename = file.path(dir_M,"git-annex/globalprep/prs_fish/v2018/tmp/temp_resample2-1.tif"), overwrite=TRUE) 
  catch_resmp <- raster(file.path(dir_M,"git-annex/globalprep/prs_fish/v2018/tmp/temp_resample2-1.tif"))
    
  
  
  # Divide Catch rasters by NPP rasters
  overlay(catch_resmp, npp, fun=function(x,y){x/y}, filename = file.path(dir_M,
                                                                         sprintf("git-annex/globalprep/prs_fish/v2018/int/%s/annual_catch_corr_%s.tif",layer,yr)
                                                                         
                                                                         ), overwrite=TRUE)
  
}

## Specify name of layer
lapply(comm_land, catch_npp_fun, layer = 'comm_landings')

```

## Extract using start and "end" pattern
* Extract string where 'Japan' is at the very end
* Extract string where 'Japan' just comes after some string following 'Walleye pollock'
```{r}

RAM <- read.csv("https://rawgit.com/OHI-Science/ohiprep_v2018/master/globalprep/fis/v2018/int/RAM_fao_ohi_rgns.csv")
RAM[which(
  
  str_detect(RAM$stocklong, "^Walleye pollock.*Japan")
  
  ),]

RAM[which(str_detect(RAM$stocklong, paste(c("^Walleye pollock.*Japan$"), collapse = "|"))),]

```


# Case When Example

1. Setup

```{r}

library(dplyr)

```

```{r}

RAM_rgns_new <- read.csv("globalprep/fis/v2018/int/RAM_fao_ohi_rgns.csv")

RAM_filt <- RAM_rgns_new %>% 
  filter(is.na(fao_id)) %>% 
  select(-RAM_area_m2)


RAM_ifelse <- RAM_filt %>%
  ## Adding FAO region id
  mutate(fao_id  = 
           ifelse(str_detect(stocklong, paste(c("Bering Sea","Gulf of Alaska$"), collapse = "|")), 61, 
                  ifelse(str_detect(stocklong, "Sicily$"), 37,
                         ifelse(str_detect(stocklong, paste(c("^Jackass morwong Western","^Jackass morwong Eastern","Redfish Eastern Australia","Blue grenadier New South Wales"), collapse = "|")), 57, NA)))) %>% 
  
  mutate(fao_id_2  = 
           ifelse(str_detect(stocklong, "Bering Sea"), 67, 
                  ifelse(str_detect(stocklong,  paste(c("^Jackass morwong Eastern","Redfish Eastern Australia","Blue grenadier New South Wales"), collapse = "|")), 81,NA))) %>% 
  
  ## Adding OHI region id
  mutate(rgn_id  = 
           ifelse(str_detect(stocklong, paste(c("Bering Sea","Gulf of Alaska$"), collapse = "|")), 163,
                  ifelse(str_detect(stocklong, "Sicily$"), 184,
                         ifelse(str_detect(stocklong, "Australia$"), 16, NA))))

DT::datatable(RAM_ifelse,rownames = F)

```

* Don't have to worry about counting number of nested parentheses - just need the `mutate` and `case_when` closing parentheses.
* More organized alignment
* Automatic if case doesn't match then doesn't do anything
```{r}

RAM_case <- RAM_filt %>% 
  ## ADDING FAO REGION ID
  mutate(fao_id = 
           case_when(
             
             str_detect(stocklong, paste(c("Bering Sea","Gulf of Alaska$"), collapse = "|")) ~ 61,
             str_detect(stocklong, "Sicily$") ~ 37,
             str_detect(stocklong, paste(c("^Jackass morwong Western","^Jackass morwong Eastern","Redfish Eastern Australia","Blue grenadier New South Wales"), collapse = "|")) ~ 57
             
             )) %>% 
  
  mutate(fao_id_2 =
           case_when(
             str_detect(stocklong, "Bering Sea") ~ 67,
             str_detect(stocklong,  paste(c("^Jackass morwong Eastern","Redfish Eastern Australia","Blue grenadier New South Wales"), collapse = "|")) ~ 81
             )) %>% 
  
  ## ADDING OHI REGION ID
  mutate(rgn_id = 
           case_when(
             
             str_detect(stocklong, paste(c("Bering Sea","Gulf of Alaska$"), collapse = "|")) ~ 163,
             str_detect(stocklong, "Sicily$") ~ 184,
             str_detect(stocklong, "Australia$") ~ 16
             
           )) 
  

DT::datatable(RAM_case,rownames = F)

```

Compare - check differences

```{r}

setdiff(RAM_case, RAM_ifelse)
setdiff(RAM_ifelse, RAM_case)

```

