---
title: "Spatial Training"
output: 
  html_document:
    toc: TRUE
    toc_float: TRUE
---

# Introduction to Spatial in R

Vector spatial data

* `sp`/`rgdal` - closely linked packages, used when working with spatial files (e.g. read/write spatial data, transform coordinate system)
* `regeos` - can merge polygons and use the buffer and intersect operations
* `sf` - can treat spatial data as a data frame and use similar approaches to manipulating data; sometimes need to transform back to shapefile format but may eventually replace `sp` and `rgeos`; can't interact with rasters though - need to convert back to shapefile

Raster spatial data

* `raster` - primarily used for raster manipulation (read/write, manipulate, transform, analyze, modeling)
* `fasterize` - converts vector data into a raster (similar to raster::rasterize, which is slower)
* `rgdal` - used for projections

[spatial R script for OHI](https://github.com/OHI-Science/ohiprep_v2018/blob/bccb633f6e147142e814eae4a84f81a572b2d369/src/R/spatial_common.R)
[rspatial on twitter](https://twitter.com/hashtag/rspatial)
[Jamie's spatial analysis tutorial](http://jafflerbach.github.io/spatial-analysis-R/)

# Projections

* Can't really preserve both direction and distance/area simultaneously (trade-offs!)
* Lat/Long is considered unprojected data; most common for global data, but not great for very localized data (e.g. 1-m resolution)
* Mollweide represents area more accurately; most common for OHI global assessments; direction not accurately represent, but not important for OHI; not commonly used in the general community (people might not be familiar)
* Data must be in same projection: specify with epsg code or proj4string

Vector spatial data

* Retrieve projection - `proj4spring(x)`
* Describe/define projection - `proj4spring(x) <- CRS("+init=epsg:28992")` OR `proj4spring(x) <- CRS("+proj=moll +ellps=WGS84")`
* If need to transform from one CRS projection to another, can use `spTransform(x, CRS("..."))`

Raster spatial data

* Retrieve - `projection(x)`
* Describe - `projection(x) <- CRS("...")`
* Transform - `projectRaster()`


# Exploring `sf`

See [Jamie's training doc](https://rawgit.com/jafflerbach/spatial-analysis-R/gh-pages/intro_to_sf.html).

## Load Packages
```{r}

# install.packages(c("sf","raster","rgdal"))
# install.packages("fasterize")

## load packages using library()
library(tidyverse) #devtools::install_github("tidyverse/ggplot2") ## if most recent vers. not in CRAN
library(sp)
library(rgdal)
library(sf)
library(raster) # allows you to print out shapefiles with a more condensed summary rather than the full-length list; dplyr and raster may have conflicts
library(fasterize)

# select <- dplyr::select ## this code will force select() to be used from the dplyr package not raster package

```

## 1. Load shapefile using `rgdal`
```{r}

setwd("../spatial-analysis-R/")

west_shp_rgdal <- rgdal::readOGR(dsn = "shapefiles", layer = "wc_regions")

## Take a look in a plot
plot(west_shp_rgdal)

## Check the class of the object is SpatialPolygonsDataFrame
west_shp_rgdal

```

## 2. Load shapefile using `sf`

* Will print out some of the data table associated with the shapefile
* With an sf object always subset for 1 column when plotting so R doesn't crash if it's huge
```{r}

setwd("../spatial-analysis-R/")
west_shp_sf <- st_read(dsn = "shapefiles", layer = "wc_regions")
west_shp_sf
plot(west_shp_sf[1])
nrow(west_shp_sf)

```

## `sf` and `dplyr`
```{r}

## remove a column using sf & dplyr
west_shp_sf %>% 
  dplyr::select(rgn, rgn_key) # geometry column is always preserved, but removes 'area_m2'

## filter
west_shp_sf %>% 
  dplyr::filter(rgn == "Oregon")

## Create new two new columns with mutate
sf_out <- west_shp_sf %>% 
  mutate(rgn_id = c(1:5),
         area_km2 = area_m2/1000000)

```

## Transform spatial data in `sf`

* Currently sf_out is in lat/long format
* Let's reproject sf_out to Mollweid
```{r}

sf_trans <- st_transform(sf_out, crs = "+proj=moll +ellps=WGS84")
plot(sf_trans[1])

```

## Save simple feature
```{r}

setwd("../spatial-analysis-R/")
st_write(sf_out, "shapefiles/wc_regions_clean.shp", driver = "ESRI Shapefile", delete_layer = TRUE) # delete_layer lets you overwrite the file if it already exists and replace it

```

## Visualize with ggplot
```{r}

ggplot(sf_out) +
  geom_sf()

## Plot and fill with a discrete variable
ggplot(sf_out) +
  geom_sf(aes(fill = rgn)) + # rgn is discrete variable
  theme_minimal() +
  labs(fill = "Region") + # this changes the legend title to "Region"
  scale_fill_brewer(palette = "Spectral")

## Plot and fill with a continuous variable
ggplot(sf_out) +
  geom_sf(aes(fill = area_km2)) + # area is continuous variable 
  theme_bw() +
  labs(fill = "Area (km2)") +
  scale_fill_distiller(palette = "Blues", direction=1) #direction can be used to reverse the color palette scale; try changing value to '-1"

```

## Manipulate shapefiles

Use `st_union` to merge all polygons
```{r}

## Merge all polygons (it'll convert from sf class to sfc - no longer a data frame)
full_rgn <- st_union(sf_out)
plot(full_rgn[1])
class(full_rgn) # it is now class sfc which can't be plotted in ggplot

## Convert full_rgn back to an sf object using st_sf
full_rgn <- st_union(sf_out) %>%
            st_sf(geometry = .) # the "." calls the object that was created last, st_union(sf_out)

class(full_rgn) # now it's class sf and data.frame, which allows you to manipulate it with tidyverse tools

 ggplot(full_rgn) +
   geom_sf()
 
```


# Exploring `raster`

## Import data
```{r}

## west coast regions shapefile and take a look at it
wc_rgns <- st_read(dsn = "shapefiles", layer = "wc_regions_clean")
plot(wc_rgns[1])

## read in a raster
r <- raster("rasters/average_annual_sst_2008.tif")
plot(r)

## Take a look at raster data with histogram
hist(r)

```

## Combine & manipulate all rasters
```{r}

## Read in all of the sst raster layers and put them into a stack
sst_files <- list.files("rasters", pattern = "average_", full.names=T)

sstStack <- stack(sst_files)
plot(sstStack)

## Get the average SST in Celcius
## function averages SST and converts to C
myFUN <- function(x){
  mean(x, na.rm=T)-273.15
  }

sst_avg <- raster::calc(sstStack, fun=myFUN, filename = "rasters/sst_avg.tif") # calculates avg and saved to a file (good for large data)
plot(sst_avg)
```

## Import in NPP raster
```{r}

## Net primary productivity
npp <- raster("rasters/annual_npp.tif")
plot(npp)

## Some exploration tools you can use
#zoom(npp)
#cropped <- zoom(npp)
#click(npp)
```

```{r}
## Transform into same projection as SST data
nppProj <- raster::projectRaster(npp, crs = ("+proj=longlat"))
# nppProj <- raster::projectRaster(npp, crs = crs(sst_avg)) # other way to transform
plot(nppProj)
plot(wc_rgns, add = T) # add wc_rgns to the current plot

## Crop nppProj to sst_avg bounding coordinates
nppCrop <- raster::crop(nppProj, sst_avg)
plot(nppCrop)
plot(wc_rgns, add=T)

## Resample - divide one cell into multiple cell
npp_res <- raster::resample(nppCrop, sst_avg, method = "ngb") # nearest neighbor - keep data as close to raw data as possible (no new data is created)
npp_bil <- raster::resample(nppCrop, sst_avg, method = "bilinear") # tries to smooth across cells, averaging between original cells

## plot two figures next to each other
par(mfrow = c(1,2))
plot(npp_res); plot(npp_bil)

dev.off() # removes the par() setting

## Try using pipes to do it all together
npp_res <- npp %>%
  raster::projectRaster(crs = ("+proj=longlat")) %>% 
  raster::crop(sst_avg) %>% 
  raster::resample(sst_avg, method = "ngb", progress = "text") # progress gives you a progress bar while you're running the code

plot(npp_res)  

stack(npp_res, sst_avg)

## Reclassify values into 0 or 1 to indicate suitability for aquaculture 
## Optimal temp is between 12 and 18 C
sstPref <- reclassify(sst_avg, c(-Inf, 12, NA, # all cells between -Inf and 12, turn it into NA
                                12, 18, 1, # all cells between 12 and 18, turn it into 0
                                18, Inf, NA)) # all cells between 18 and inf, turn it into NA

plot(sstPref)

## Reclassify NPP
## NPP constraints between 2.6 and 3

nppPref <- reclassify(npp_res, c(-Inf, 2.6, NA,
                                 2.6, 3, 1,
                                 3, Inf, NA))

plot(nppPref)

## Combine rasters using overlay() function
## Specify how you want R to overlay the raster. Here we want desirable cells to have a value of 1
cells <- raster::overlay(sstPref, nppPref, fun=function(x,y){x*y})
plot(cells, col="lightblue")

## Mask - remove all cells that fall outside the US EEZ
cellsEEZ <- mask(cells, wc_rgns) # removes parts of cells that lies outside of wc_rgns
plot(cellsEEZ)

## Rasterize wc_rgns (convert from simple features to raster)
wc_rgns # print to see the data table (it's a sf so you can see this info, once rasterized you can't)
wcZones <- fasterize(wc_rgns, cellsEEZ, field = "rgn_id") # cellsEEZ used as a raster template for the raster output of wc_rgns
plot(wcZones)

## Zonal stats
suit_area <- raster::zonal(cellsEEZ, wcZones, fun = "sum") # wcZones is the raster with codes for the 5 zones, add up the values in cellsEEZ 
suit_area # adds up cells in each zone (e.g. Washington)

```

## Additional Visualizing and Calcs
```{r}

## Plot the suitable aquaculture areas and US west coast zones next to each other
par(mfrow = c(1, 2))
plot(cellsEEZ, col='lightblue', main='Suitable Aquaculture Areas');plot(wcZones, main='US West Coast zones')

## Create a new raster with area values in the data
cellsArea = area(cellsEEZ, na.rm=T)
plot(cellsArea, main='Cell area (km2)') 

```


```{r}
sf_trans <- st_transform(sf_out, crs = "+proj=moll +ellps=WGS84")
wcZones
projection(wcZones) <- CRS("+proj=moll +ellps=WGS84")
wcZones
plot(wcZones)
sf_trans
plot(sf_trans)

new <- fasterize(sf_trans, wcZones, field = "rgn_id")
plot(new)

```



