---
title: "Intro to Strings"
author: "Iwen Su"
date: "3/16/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load packages

```{r}

library(tidyverse)
library(stringr)

```

## Exploring the `stringr` functions

```{r}

# create vector of strings
x <- c("why", "video", "cross", "extra", "deal", "authority")

# gives the number of characters in each string in the vector
str_length(x) 

# combine the vector into a single string using `collapse` and indicating what string to insert between each string - default is collapse = ""
str_c(x, collapse = ", ")
str_c("why","you", sep = ", ") # does the same thing
str_c(x, sep = ", ") # doesn't collapse into one string
str_c(x, x, sep = ",") # collapses each [i]th of the first and second string argument 

# only print out the first three letters of each string
str_sub(x, 1, 3)
# only print out the letters in each string from the second to the third letter
str_sub(x, 2, 3)

# subset the vector and only keep strings that contain vowels aeiou ("why" is removed)
str_subset(x, "[aeiou]")
# count the number of vowels in each string
str_count(x, "[aeiou]")


```

## 7 Main Verbs that deal with Patterns

```{r}

## Here the pattern is "[aeiou]": presence of vowels

# returns a TRUE/FALSE detection of the pattern
str_detect(x, "[aeiou]")

# count number of the pattern
str_count(x, "[aeiou]")

# extracts strings containing the pattern
str_subset(x, "[aeiou]")

# gives position of the match
str_locate(x, "[aeiou]")

# extracts the pattern in each string (only prints first match, not if there are multiple)
str_extract(x, "[aeiou]")

# extracts characters on either side of the vowel
str_match(x, "(.)[aeiou](.)")

# replaces the pattern with a diff string/text
str_replace(x, "[aeiou]", "?")

# splits up string into multiple pieces based on a specified string
str_split(c("a,b", "c,d,e"), ",")
str_split("hey-you-what-is-going-on","-")
str_split("testabtestabhiabwhoab","ab")

```

## R 4 Data Science - Chapter 14: Strings

```{r}

# use \ to 'escape' a symbol and indicate it as part of the string
double_quote <- "\"" # printed variable is not the string (includes escapes)
backslash <- "\\" 

writeLines(double_quote) # use this function to view string

# see help page and all special characters: 
?"`"

# there is code for certain symbols
x <- "\u00b5"
writeLines(x)

```

## Exploring strings

```{r}

x <- c("abc", NA)
str_c("|-", x, "-|")
str_c("|-", str_replace_na(x), "-|")

name <- "Hadley"
time_of_day <- "morning"
birthday <- FALSE

str_c(
  "Good ", time_of_day, " ", name,
  if (birthday) " and HAPPY BIRTHDAY",
  "."
)

```

## Matching patterns

```{r}

x <- c("apple", "banana", "pear")
str_view(x, "an")
str_view(x, ".a.")

# To create the regular expression, we need \\
dot <- "\\."

# But the expression itself only contains one:
writeLines(dot)
#> \.

# And this tells R to look for an explicit .
str_view(c("abc", "a.c", "bef"), "a\\.c")

x <- "a\\b"
writeLines(x)
#> a\b

str_view(x, "\\\\")

```

## Matching patterns: anchor to start or end

* `^` to match start of string
* `$` to match end of string

```{r}

x <- c("apple", "banana", "pear")
str_view(x, "^a")
str_view(x, "a$")

x <- c("apple pie", "apple", "apple cake")
str_view(x, "^apple$")

```

## Character classes and alternatives

* `\d`: matches any digit.
* `\s`: matches any whitespace (e.g. space, tab, newline).
* `[abc]`: matches a, b, or c.
* `[^abc]`: matches anything except a, b, or c.

```{r}

str_view(c("grey", "gray"), "gr(e|a)y")

```

## Detect matches

* `str_detect()`
* `str_count()`

```{r}

x <- c("apple", "banana", "pear")
str_detect(x, "e")

a <- "bad"
b <- "bad"
identical(a,b)

str_subset(words, "x$")

df <- tibble(
  word = words, 
  i = seq_along(word)
)

# filter for words that match "x$"
df %>% 
  filter(str_detect(words, "x$"))

df %>% 
  mutate(
    vowels = str_count(word, "[aeiou]"),
    consonants = str_count(word, "[^aeiou]")
  )

```

## Match and extract 
```{r}

colours <- c("red", "orange", "yellow", "green", "blue", "purple")
colour_match <- str_c(colours, collapse = "|")

# subset for sentences with a color match
has_colour <- str_subset(sentences, colour_match)

# which colors matched? - only extracts first match
matches <- str_extract(has_colour, colour_match)
head(matches)

more <- sentences[str_count(sentences, colour_match) > 1]
str_extract_all(more, colour_match)

```

