---
title: "Data Transformation Training"
author: "Iwen Su"
date: "3/8/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Intro to Data Transformation

[R-4-DataScience][http://r4ds.had.co.nz/transform.html]

### Load packages

```{r}

library(nycflights13) # install.packages("nycflights13")
library(tidyverse)

```

### Grab data

```{r}

flights <- nycflights13::flights

```

### Key dplyr functions:

`filter()`, `arrange()`, `select()`, `mutate()`, `summarize()`, & `group_by()`

#### `filter()`

```{r}

# Using the `filter()` function
# To print out results AND save to variable wrap with parentheses
(jan1 <- flights %>% 
  filter(month == 1, day == 1)
)

novdec <- flights %>% 
  filter(month == 11 | month == 12)
head(novdec)
tail(novdec)

nov_dec <- flights %>% 
  filter(month %in% c(11,12))

delay <- flights %>% 
  filter(!(arr_delay > 120 | dep_delay >120))

delay <- flights %>%  
  filter(arr_delay <= 120, dep_delay <= 120)

# using filter while preserving NA values
df <- tibble(x = c(1,NA,3))
df %>% 
  filter(x > 1 | is.na(x))

# Exercises
flights %>% 
  filter(arr_delay >= 2)
flights %>% 
  filter(dest %in% c("IAH","HOU"))
flights %>% 
  filter(carrier %in% c("UA","AA","DL"))
flights %>% 
  filter(month %in% c(7,8,9))
flights %>% 
  filter(dep_delay <= 0 & arr_delay > 2)
flights %>% 
  filter(dep_delay >= 1 & arr_delay <= -30)
flights %>% 
  filter(dep_time < 600 & dep_time > 000)

# use shortcut `between()`!
flights %>% 
  filter(between(dep_time, 000, 600))


```

#### Using `arrange()`

```{r}

# first arrange by year then by month then by day
flights %>% 
  arrange(year, month, day)

# arrange in descending order
flights %>% 
  arrange(desc(arr_delay)) %>% 
  tail()

# NA values get sorted to the end, always

# Exercises
# arrange NA values to the start
flights %>% 
  arrange(desc(is.na(dep_time)))

# sort flights for most delayed flights
flights %>%
  arrange(desc(dep_delay))

# sort flights for those that left the earliest
flights %>%
  arrange(dep_delay)

# fastest flight
flights %>% 
  arrange(hour, minute)

# shortest flight by distance was 17 miles, longest is 4983 miles
flights %>% 
  arrange(distance) %>% 
  tail()

```

#### Selecting columns with `select()`

Learning about helper functions!! `starts_with`, `ends_with`, `contains`, `matches`, `num_range`

```{r}

# different ways to select

flights %>% 
  select(year, month, day)

flights %>% 
  select(year:day)

# all but year, month and day columns
flights %>% 
  select(-(year:day))

flights %>% 
  rename(tail_num = tailnum) # new column name = original column name

# rearrange some columns to the front
flights %>% 
  select(time_hour, air_time, everything())

# Exercises
flights %>% 
  select(arr_time, arr_delay, dep_time, dep_delay)
flights %>% 
  select(dep_time:arr_delay, -sched_dep_time, -sched_arr_time)

flights %>% 
  select(dep_time, dep_time)

# Unsure what `one_of` is supposed to do
vars <- c("year", "month", "day", "dep_delay", "arr_delay")
flights %>% 
  select(one_of(vars))

# default is to ignore case, but if you change value to FALSE then it is case-sensitive
flights %>% 
  select(contains("TIME", ignore.case = FALSE))


```

#### Add new variables with mutate()

```{r}

flights_sml <- flights %>% 
  select(year:day, 
         ends_with("delay"),
         distance,
         air_time)

# create new columns with mutate and refer to newly created columns
flights_sml %>% 
  mutate(gain = arr_delay - dep_delay,
         hours = air_time/60,
         gain_per_hour = gain / hours
         )

# create new columns and remove the rest using `transmute`
flights %>% 
  transmute(gain = arr_delay - dep_delay,
            hours = air_time / 60, 
            gain_per_hour = gain/hours
            )

```

### Miscellaneous 

```{r}

# using near()
near(sqrt(2)^2,2)
near(1/49*49, 1)

# using as.na() to ask if a value is missing
x <- c(1,2,3,4,NA)
is.na(x)

```


# LOOKS LIKE I LOST ALL THE PROGRESS AFTER MISCELLANEOUS ^ :(




### 5.2.3 Missing values

```{r}

library(nycflights13)

# create tibble with single column
df <- tibble(x = c(1, NA, 3, NA), y = c(2, NA, 4, 6))

# filtering with base R
# filter(df, x > 1 | is.na(x))
# filter with dplyr
df %>% 
  filter(is.na(x) | x > 1)
df %>% 
  filter(is.na(x), x > 1) # this won't work because you can't simultaneously be an NA value in the x column and greater than 1

# both of these read: filter for values in column x that is greater than or equal to 1 AND the value in column y must simultaneously equal to 4
df %>% 
  filter(x >= 1, y == 4) 
df %>% 
  filter(x >= 1 & y == 4)

# Filter for rows that don't have NAs in both x and y column
df %>% 
  filter(!(is.na(x) & is.na(y)))
df %>% 
  filter(!is.na(x) | !is.na(y)) # down for rows that don't have NAs in x or don't have NAs in y, but will grab rows where y has a value but x is an NA or vice versa

# select rows where either x is greater than 1 or y is greater than 4
df %>% 
  filter(x > 1 | y > 4)
df %>% 
  filter(is.na(x) | is.na(y))



# Both of these read: filter for rows where the arrive delay is less than or equal to 120 and the departure delay is less than or equal to 120. ???
# any columns where arrival delay is greater than 120 even if depart delay is < 120 remove. any columns where departure delay is greater than 120 even if arrival delay is less than 120. remove!
filter(flights, !(arr_delay > 120 | dep_delay > 120))
filter(flights, arr_delay <= 120, dep_delay <= 120)

```





















